---
title: "Externe Daten"
author: "A. Paul"
date: "September 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(fars)
library(readr)
library(dplyr)
```

## External Dataset

The statistics about the accidents are collected and published by the *US National Highway Traffic Safety Administration's Fatality Analysis Reporting System (NHTSF)*, which is a nationwide census providing the American public yearly data regarding fatal injuries suffered in motor vehicle traffic crashes.

More information and the data files can be found on the website 

  *@source*  https://www.nhtsa.gov/Data/Fatality-Analysis-Reporting-System-(FARS).

There are three data file (‘accident_2013’, ‘accident_2014’ and ‘accident_2015’) in the directory ~/inst/extdata.

They are used for tests, for examples and for show how to work with an dataset.


## @Format 

The fields of the csv files are shown here.   

```{r format}
df <- fars_read( filename = "accident_2015.csv.bz2" )
str(df)
```

## Example

Let's have a look at an output of the first 10 table entries for some fields.

```{r example}
knitr::kable( head(df, 10) )
```
