---
title: "Vignette for Package fars"
author: "A. Paul" 
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introducing fars}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set( echo = TRUE )
```

```{r libs, include=FALSE}
library(fars)
library(readr)
library(dplyr)
library(magrittr)
library(tidyr)
library(bindrcpp)
library(maps)
library(graphics)
```

```{r coll, include=FALSE}
knitr::opts_chunk$set(comment = "#>", collapse = TRUE)
```

## Introduction to package "fars"

The `"fars"` package is used to evaluate the reported accidents in the United States of America.

The statistics are collected and published by the *US National Highway Traffic Safety Administration's Fatality Analysis Reporting System (NHTSF)*, which is a nationwide census providing the American public yearly data regarding fatal injuries suffered in motor vehicle traffic crashes.

More information and the data files can be found on the website 
https://www.nhtsa.gov/Data/Fatality-Analysis-Reporting-System-(FARS).

The package supports the following evaluations by means of functions:

* a tabular analysis of accidents per month and year, and
* a graphical representation of the accidents per US state per year

The following functions of the "fars" package are supported and are described:

* Reading a file using the function `fars_read()`
* Provision of a dataframe using `fars_summarize_years()`
* Creation of a plot using `fars_map_state()`

## Prerequisites  

1. The following packages are used:

* library(fars)
* library(readr)
* library(dplyr)
* library(magrittr)
* library(tidyr)
* library(bindrcpp)
* library(maps)
* library(graphics)

2. Download csv files from Website of NHTSF:
The prerequisite for performing the evaluations is that the csv files ("accident_yyyy.csv.bz2") are downloaded with the accident data and are located in the current directory.

    *Website*  https://www.nhtsa.gov/Data/Fatality-Analysis-Reporting-System-(FARS)

## Read dataset with function fars_read() 

The function `fars_read (filename)` is used to read a file (eg "accident_2015.csv.bz2") with accidents of the NHTSF.
The argument of the function is the filename and the return is the tibble (or data frame). 


###### Excample

```{r, farsread}
tf <- "accident_2015.csv.bz2"
df <- fars_read( filename = tf )
str(df)
```

Let's have a look at an output of the first 10 table entries for some fields.

```{r, fars_read_tab}
## selection of some fields 
df_sel <- select( df, STATE, PERSONS, COUNTY, CITY, DAY, MONTH, YEAR, SP_JUR, FATALS )
knitr::kable( head(df_sel, 10) )
```


## Prepare fars data with function fars_summarize_years()

The function `fars_summarize_years(years)` is used to create a tabular representation of the accidents per month and year. The argument of the function is a vector with the years to be considered for the creation and the return is the tibble (or data frame).

###### Excample

```{r, fars_summ}
yyyys <- c(2014, 2015)
df_summ <- fars_summarize_years( years = yyyys )
str(df_summ)
```

Here is the output for the year 2015. 

```{r, fars_summ_tab}
knitr::kable(head(df_summ, 13))
```


## Plot the distributation of accidents in an US state with function fars_map_state()

The function `fars_summarize_years (state.num, year)` is used to create a plot with the distribution of all accidents in one year on the map of a US state.

###### Excample

```{r, fig.height=5, fig.width=7}
state_no <- 01
par( mfrow = c(1, 3) )
## title(main="Accidents in the US State Alabama")
yyyy <- 2013
fars_map_state( state.num=state_no, year=yyyy )
title(main="Alabama: Accidents in 2013")
yyyy <- 2014
fars_map_state( state.num=state_no, year=yyyy )
title(main="Alabama: Accidents in 2014")
yyyy <- 2015
fars_map_state( state.num=state_no, year=yyyy )
title(main="Alabama: Accidents in 2015")
```
